Model based reinforcement learning at FERMI FEL

### 2020-11-16 main.tex
txs auto checkin

%\tableofcontents
\section{Introduction and Motivation}
In particle accelerators one main goal is to provide stable and reproducible performance. In order to achieve this, a number of control problems have to be considered simultaneously. Especially if there is no way to model the physics, one might use optimisation techniques as e.g. derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes. Another promising way is to apply reinforcement learning which has several advantages over optimisation methods:
\begin{itemize}
    \item Covers a larger class of problems.
    \item Solves the problem not always completely from the beginning as an DFO.
    \item Ability to use existing data.
    \item The underlying structure of the problem might be deduced.
\end{itemize}
RL optimises a sequence of decisions. 
In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste. Three algorithms were applied successfully showing different advantages, which will e discussed in the following.

@@

\section{Uncertainty aware Dyna-style reinforcement learning}
The original Dyna algorithm \cite{DYNA} is modified here in several aspects. Generally, Dyna style algorithms denote algorithms, where a MFRL algorithm is trained on purely synthetic data from an approximate dynamics model or on a mixture of synthetic and real data. We use only synthetic data to reduce the interaction with the real environment to a minimum.

>>>>

%\tableofcontents
\section{Introduction and Motivation}
In particle accelerators one main goal is to provide stable and reproducible performance. In order to achieve this, a number of control problems have to be considered simultaneously. Especially if there is no way to model the physics, one might use optimisation techniques as e.g. derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes. Another promising way is to apply reinforcement learning which has several advantages over optimisation methods:
\begin{itemize}
    \item Covers a larger class of problems.
    \item Solves the problem not always completely from the beginning as an DFO.
    \item Ability to use existing data.
    \item The underlying structure of the problem might be deduced.
\end{itemize}
RL optimises a sequence of decisions. On critical aspect using RL is the number of iteration to train a controller.
In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste. Three algorithms were applied successfully showing different advantages, which will e discussed in the following.

@@

\section{Uncertainty aware Dyna-style reinforcement learning}
The original Dyna algorithm \cite{DYNA} is modified here in several aspects. Generally, Dyna style algorithms denote algorithms, where a MFRL algorithm is trained on purely synthetic data from an approximate dynamics model or on a mixture of synthetic and real data. We use only synthetic data to reduce the interaction with the real environment to a minimum.
An overview of the used method is shown in \cref{fig:MBRL_overview}. At the beginning the data is collected or read in from a pre-collection. An uncertainty aware model is trained, using anchored ensembles on the data, which allows to take the allegorical (measurement errors) as well as the epistemic uncertainty (lack of data) into account. Subsequently a MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data, by taking the uncertainty into account. After a defined number of training steps the controller is testes on each model individually. If there is no improvement in a specific ratio<1 of models, the controller is tested on the real environment.

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/MBRL_overview}
	\caption{A schematic overview of the AE-Dyna approach used in this paper.}
	\label{fig:MBRL_overview}
\end{figure}
